### 2015-04-26 paper.tex

\documentclass[envcountsame]{cls/cccg15}
\usepackage{graphicx}

\usepackage{amssymb,amsmath}
\usepackage[mathscr]{euscript}

\usepackage{caption,subcaption}
\usepackage{booktabs,multirow}

\usepackage{algorithm}
\usepackage[noend]{algorithmic}

\usepackage{cleveref}

%------------------------ Paper Notations -----------------------------

\newcommand{\rc}{r}
\newcommand{\cp}{c_p}
\newcommand{\dz}{(d + 1)(z + 1)}
\newcommand{\Buffer}{\ensuremath{\text{Buffer}}}

\renewcommand{\O}{\ensuremath{{O}}} 
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\nd}{\nobreakdash-}

%----------------------- Page Style -----------------------------------------

\renewcommand{\arraystretch}{1.3}

%----------------------- Environments --------------------------

\newtheorem{problem}{Problem}
\newtheorem{definition}{Definition}

%------------------------ Algorithm -----------------------------

\newcommand{\textproc}{\textsc}
\newcommand{\Call}[2]{\textsc{#1}(#2)}
\renewcommand{\algorithmicforall}{{\bf for each}}

%-------------------------- Notations ------------------------------------

\newcommand{\IN}{\ensuremath{\mathbb{N}}} 
\newcommand{\IZ}{\ensuremath{\mathbb{Z}}} 
\newcommand{\IQ}{\ensuremath{\mathbb{Q}}} 
\newcommand{\IR}{\ensuremath{\mathbb{R}}} 
\newcommand{\IC}{\ensuremath{\mathbb{C}}}

\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\seq}[1]{\left< #1 \right>}
\newcommand{\ceil}[1]{\left\lceil{#1}\right\rceil}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
\newcommand{\card}[1]{\left|{#1}\right|}
\newcommand{\len}[1]{\|{#1}\|}
\newcommand{\setcomp}[1]{\overline{#1}}
\newcommand{\provided}{\,|\,}

\newcommand{\poly}{\mathop{\mathrm{poly}}}
\newcommand{\polylog}{\mathop{\mathrm{\scriptsize polylog}}}
\newcommand{\lcm}{\mathop{\mathrm{lcm}}}

\newcommand{\lee}{\leqslant}
\newcommand{\gee}{\geqslant}
\renewcommand{\leq}{\lee}
\renewcommand{\le}{\lee}
\renewcommand{\geq}{\gee}
\renewcommand{\ge}{\gee}
\newcommand{\lei}{\prec}

\newcommand{\OPT}{\ensuremath{\mathop{\mathrm{OPT}}}}
\newcommand{\APX}{\ensuremath{\text{\sc APX}}}
\newcommand{\opt}{\ensuremath{\mbox{\tt opt}}}
\newcommand{\apx}{\ensuremath{\mbox{\tt apx}}}

\newcommand{\eps}{\varepsilon}
\newcommand{\bslash}{\!\setminus\!}
\newcommand{\bigOmega}{{\rm\Omega}}
\newcommand{\etal}{{\em et~al.\/}}
\newcommand{\REM}[1]{}
\newcommand{\qed}{}
\newcommand{\backspace}{\vspace{-3em}\\}

\newcommand{\linprog}[6]{
  \begin{alignat}{2}
    \text{#1}  \quad & #2\\
    \text{s.t.}  \quad & #3 & \qquad #4 \notag\\
                     & #5 & \qquad #6 \notag
  \end{alignat}}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

%----------------------- Title -----------------------------------------

\title{A Streaming Algorithm for 2-Center with Outliers in High Dimensions}
%\title{Computing 2-Center with Outliers on High-Dimensional Data Streams}
\author{Behnam~Hatami\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt hatami@ce.sharif.edu}}
	\and 
	Hamid~Zarrabi-Zadeh\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt zarrabi@sharif.edu}}
}

%------------------------------ Text -------------------------------------

\begin{document}

\maketitle
\pagestyle{plain}

% ---------------------- Abstract --------------------------------------

\begin{abstract}
We study the Euclidean 2-center problem with outliers in the data stream model. 
Given a stream of points in arbitrary $d$ dimensions, the goal is to find two congruent balls 
of minimum radius covering all but $z$ points. 
We provide a $(1.8+\eps)$-approximation streaming algorithm for the problem, 
improving upon the previous $(4 + \eps)$-approximation algorithm available for the problem.
The space complexity and update time of our algorithm is $\poly(d, z, {1 \over \eps})$,
independent of the size of the stream.
\end{abstract}

% ---------------------- Introduction ---------------------------------

\section{Introduction}
The $k$-center problem---covering a set of points 
using $k$ congruent balls of minimum radius---is a fundamental problem,
arising in many applications 
such as data mining, unsupervised learning, statistics, and image processing.
In real-world applications, where input data is often noisy, 
it is very important to consider outliers, 
as even a small number of outliers can greatly affect the quality of the solution.
The $k$-center problem is particularly very sensitive to outliers,
and even a constant number of outliers can increase the radius of the $k$-center unboundedly.
Therefore, it is natural to consider the following generalization of
the the $k$-center problem: % which we call \emph{$k$-center with $z$ outliers}:
given a set $P$ of $n$ points in arbitrary $d$ dimensions, 
find $k$ congruent balls of minimum radius 
to cover at least $n - z$ points of $P$.
%One of $k$-center problem variants is $k$-center with $z$-outlier which finds $k$ congruent balls with minimum radius covering all but $z$ points of a point set.

In this paper, we focus on the \emph{data stream model}
where only a single pass over the input data is allowed,
and we have only a limited amount of working space available.
%which is typically sublinear in the input size.
This model is specially useful for processing large data sets,
as it does not need the entire data set to be stored in memory.

%In this model, input data arrives as a sequence over time, 
%%only one pass over the input data is allowed, 
%and we have only a limited amount of working space available, 
%which is typically sublinear in the input size.
%and thus, we cannot keep the entire data set in memory.
%This makes designing streaming algorithms very challenging,
%%as we can have only one pass over the input data, 
%%and must decide on the fly which data to keep in our limited working space,
%as we cannot keep the entire data set in memory,
%and must be still able to produce a solution comparable to 
%that of an offline algorithm that has unlimited access to the whole data set.

The Euclidean $k$-center problem has been extensively studied in the literature.
%There is a considerable amount of work on the geometric $k$-center problem.
If $k$ is part of the input,
the problem in known to be NP-hard in two or more dimensions~\cite{fowler1981optimal},
and is even hard to approximate to within a factor close to 2, 
unless P$\,=\,$NP~\cite{feder1988optimal}.
Factor-2 approximation algorithms are available for the problem 
in any dimension~\cite{gonzalez1985clustering,feder1988optimal}.
%The $k$-center problem without outliers (i.e., $z=0$) in general (Euclidean) space can not be approximated with a smaller factor than $2$ ($1.822$), unless $\mathbf{P}=\mathbf{NP}$~\cite{bern1996approximation}.
%Gonzalez~\cite{gonzalez1985clustering} showed a $2$-approximation algorithm for general case of the problem.
For small $k$ and $d$, %there are more efficient algorithms.
better solutions are available.
The 1-center problem in fixed dimensions is known to be LP-type 
and can be solved in $O(n)$ time~\cite{chazelle1996linear}.
For 2-center in the plane, the current best algorithm
runs in $\O(n \log^2 n \log^2 \log n)$ time~\cite{chan1999more}.

For $k$-center with outliers, Charikar~\etal~\cite{charikar2001algorithms} 
gave the first algorithm with an approximation factor of $3$, 
that works in any dimension.
%which is improvable to 3, if all candidate center points can be enumerated.
%depending on whether the candidate center points can be enumerated or not, respectively.
Better results are known for small $k$ in the plane.
The 1-center problem with $z$ outliers in the plane can be solved 
in $\O(n \log n + z^{3} n^{\eps})$ time, for any $\eps > 0$, 
using Matou{\v{s}}ek's framework~\cite{matouvsek1995geometric}.
Agarwal~\cite{agarwal2008efficient} gave a randomized 
$\O(n z^{7} \log^3 z)$-time algorithm for 2-center with $z$ outliers in the plane.

In the streaming model, McCutchen~\etal~\cite{mccutchen2008streaming} 
and Guha~\cite{guha2009tight} presented algorithms to maintain $(2+\eps)$-approximation 
to the $k$-center problem in %in any metric space using 
$\O(\frac{kd}{\eps} \log \frac{1}{\eps})$ space.
For $k=1$, %better algorithms are known. % in high dimensions.
%a $1.5$-approximation in high dimensions was presented by Zarrabi-Zadeh and Chan~\cite{zarrabi2006simple}.
a factor-$((1+\sqrt{3})/{2})$ approximation 
was presented by Agarwal and Sharathkumar~\cite{agarwal2010streaming} 
in high dimensions, using $O(d)$ space. 
The approximation factor was later improved 
%to $(1+\sqrt{3})/{2}$ by Agarwal and Sharathkumar~\cite{agarwal2010streaming} 
to $1.22$ by Chan and Pathak~\cite{chan2014streaming}.
%If the dimension is fixed, a $(1 + \eps)$-approximation
%can be maintained in $\O({k}/{\eps^{((d-1)/2)}})$ space 
%using the notion of core-sets~\cite{zarrabi2008core}., 
%Ahn~\etal~\cite{ahn2014computing} gave a $(2+\eps)$-approximation algorithm using $\O(\frac{d}{\eps})$ space and update time, which was improved for Euclidean spaces by 
For $k=2$, Kim~\etal~\cite{kim2014improved} 
have recently obtained a $(1.8+\eps)$-approximation  
using $\O(\frac{d}{\eps})$ space and update time.

For $k$-center with $z$ outliers in the streaming model, 
McCutchen~\etal~\cite{mccutchen2008streaming} 
%(using the idea in Charikar~\etal~\cite{charikar2001algorithms} as subroutine) 
gave a $(4+\eps)$-approximation 
%(or $(3+\eps)$-approximation, if all center points can be enumerated) 
algorithm using $\O(\frac{zk}{\eps})$ space.
When dimension is fixed, a $(1 + \eps)$-approximation to 1-center with outliers
can be maintained in $\O({z}/{\eps^{((d-1)/2)}})$ space 
using the notion of robust $\eps$-kernels~\cite{agarwal2007space, zarrabi2011almost}.
For 1-center with outliers in high dimensions, 
Zarrabi-Zadeh and Mukhopadhyay~\cite{zarrabi2009streaming} 
gave a $(\sqrt{2}\alpha)$-approximation, 
where $\alpha$ is the approximation factor of the underlying algorithm for maintaining 1-center. 
Combined with the $1.22$-approximation algorithm of Chan and Pathak~\cite{chan2014streaming},
it yields an approximation factor of $(\sqrt{2} \times 1.22) < 1.73$ 
using $O(d^3z)$ space and $\poly(d,z)$ update time.
%In fixed dimensions, there is a streaming $\O(\frac{z}{\eps^{\O(d)}})$-space algorithm~\cite{agarwal2007space, zarrabi2011almost}.

\begin{table}[t]
\centering
\begin{tabular}{|c|c|c|}
\hline
\multirow{2}{*}{\begin{tabular}[c]{@{}c@{}} \ \  Problem \ \ \end{tabular}} & \multicolumn{2}{c|}{Approximation Factor} \\ \cline{2-3} 
 & Without Outliers & With Outliers \\ \hline
1-Center & 1.22~\cite{chan2014streaming} & 1.73~\cite{zarrabi2009streaming} \\ \hline
2-Center & $1.8 + \eps$~\cite{kim2014improved} & \textbf{$1.8 + \eps$} [Here]  \\ \hline
$k$-Center & $2 + \eps$~\cite{guha2009tight,mccutchen2008streaming} & $4 + \eps$~\cite{mccutchen2008streaming}  \\ \hline
\end{tabular}
\caption {Summary of the streaming algorithms 
for $k$-center with and without outliers in high dimensions.}
\label{table:summary}
\end{table}

\paragraph{Our result}
In this paper, we study the 2-center problem with outliers in high dimensions.
We present a streaming algorithm for the problem that achieves
an approximation factor of $1.8 + \eps$, for any $\eps > 1$,
%using $O(d^3z/\eps)$ space.
using $\poly(d, z, {1 \over \eps})$ space and update time.
This improves the current streaming algorithm available for the problem
which has an approximation factor of $4+\eps$.
%Our work not only improves the best previous algorithm available for the case of 2-center,
%but is also counterintuitive in the sense that its 
The approximation factor of our algorithm matches
that of the best streaming algorithm for the 2-center problem without outliers.
This is somewhat surprising, considering that 
the best current approximation factors for streaming $k$-center with and without outliers 
differ by a factor of $\sqrt{2}$ for $k=1$,
and differ by a factor of 2  for general~$k$. %the 1-center problem.
See \Cref{table:summary} for a comparison.

To obtain our result, we have used a combination of several ideas
including parallelization construction, far/close ball separation,
and keeping a lower bound on the radius and distance of the optimal balls.
Our algorithm also utilizes the ideas used in~\cite{kim2014improved} 
for the 2-center problem with no outliers.
Nevertheless, our algorithm is much detailed and uses many new ingredients.
Indeed, handling outliers makes the $k$-center problem much harder,
as we not only need to find clusters with minimum radius, 
but we also need to decide which subset of points to cluster.
This is in particular more challenging in the streaming model,
where we only have a single pass over the input, and we must decide on the fly
which point can be an outlier, and which one can be safely dropped as a non-outlier point.

\REM{
The rest of this paper is organized as follows.
In Section~\ref{sec:1-center}, we give a simple $2$-approximation algorithm for 
the 1-center problem with outliers that uses $\O(z^2 + zd)$ space.
In the next section, we show a $(1.8 + \eps)$-approximation streaming algorithm for 2-center with $z$ outliers in high dimensions. 
%In Section~\ref{subsec:smaller} we run $z + 1$ parallel instances of the algorithm in order to find the first non-outlier point.
}

% ---------------------------- Preliminaries -------------------------------

\section{Preliminaries}
\label{sec:pre}
Let $B(c,r)$ denote a ball of radius $r$ centered at $c$. 
Given two points $p$ and $q$, we denote by $pq$ the straight line segment between $p$ and $q$,
and denote by $\len{pq}$ the length of this segment. 
Given two balls $B(c,r)$ and $B'(c',r')$, 
we define 
$\delta(B, B') = \max \set{0, \len{cc'}-r-r'}$
to be the \emph{distance} between $B$ and $B'$.
Two congruent balls $B_1$ and $B_2$ with radius $r$ are said to be \emph{$\alpha$-separated}, 
if $\delta(B_1, B_2) \geq \alpha r$.

Given an $n$-point set $P$ in $d$-dimensions,
a point $c \in \IR^d$ is called a \emph{centerpoint} of $P$,
if any halfspace containing $c$ contains at least $\ceil{{n}/({d + 1})}$ points of $P$. 
%In other words, any halfspace (or convex set) that avoids a centerpoint can contain at most $\floor{\frac{dn}{d + 1}}$ points of $P$. 
It is well-known that any finite set of points in $d$-dimensional space 
has a centerpoint~\cite{danzer1963helly}. 
The following observation is a corollary of this fact.
%~\cite{edelsbrunner1987algorithms}.

\begin{obs}
\label{obs:omitting-centerpoint}
	Given a set $P$ of $k(d+1)$ points in $d$-dimensional space, 
	there is a point $c$, not necessarily in $P$, 
	such that any convex object that does not cover $c$, 
	leaves at least $k$ points of $P$ uncovered.
\end{obs}

Given a point set $P$ and a point $p \in P$,
the \emph{$k$-furthest point from $p$}
is the $k$-th point of $P$ when points are sorted
in their decreasing distance from~$p$.

We assume the standard word-RAM model of computation. 
Each coordinate value takes a unit of space.
Thus, a point in $d$ dimensions takes $\O(d)$ space,
and basic operations on the points take $\O(d)$ time.

%-------------------------- 1-center ---------------------------------

\section{A 2-Approximation for 1-Center with Outliers}
\label{sec:1-center}

To warm up, we present a simple $2$-approximation streaming algorithm
for the 1-center problem with outliers.
It utilized a parallelization technique that 
will be used extensively in the rest of the paper.
The pseudocode is provided in \Cref{alg:1-center}.
The algorithm receives as input a stream of points, $P$,
the number of outliers, $z$, 
and a special point $c \in P$ which is guaranteed to be 
a non-outlier in the optimal solution.
It returns a ball $B$ covering all but at most $z$ points of $P$.
%We will show later how to remove this assumption.

%\begin{algorithm}
%\caption{\sc 1-Center$(P, z)$} 
%\algsetup{indent=1.5em}
%\label{alg:1-center}
%\begin{algorithmic}[1]
%
%	%\STATE Let $p_i$ denote the elements of $P$, for $i \in \set{1, \dots, n}$.
%	\STATE let $p_1, \ldots, p_n$ be the points of $P$
%	\FOR{$i = 1, \dots, z+1$}
%		%\STATE move $p_i$ to the front of $P$ %is a non-outlier point in optimal solution.
%		\STATE $B_i \gets$ \Call{1-Center-Special}{$P, z,p_i$}
%	\ENDFOR
%	%\STATE $i^* \gets \argmin_{i=1}^{z+1} r_i$
%	\RETURN smallest ball among $\set{B_1, \ldots, B_{z+1}}$
%
%\end{algorithmic}
%\end{algorithm}

\begin{algorithm}
\caption{\sc 1-Center$(P, z, c)$} 
\algsetup{indent=1.5em}
\label{alg:1-center}
\begin{algorithmic}[1]
	%\STATE {$\triangleright$ $c$ is a non-outlier point in the optimal solution}
	\STATE $B \gets B(c, 0)$
	\STATE $Q \gets \emptyset$
	\FORALL {$p$ in $P$}
		\IF {$p \notin B$}
			\STATE insert $p$ into $Q$
			\IF {$|Q| = z + 1$}
				\STATE $q$ $\gets$ closest point to $c$ in $Q$
				\STATE remove $q$ from $Q$
				\STATE $B \gets B(c, \len{cq})$
			\ENDIF
		\ENDIF
	\ENDFOR
	\RETURN {B}
\end{algorithmic}
\end{algorithm}

%The function \textproc{1-Center-Special} in \Cref{alg:1-center-special}, 
%takes as input a point $c$ which is guaranteed to be a non-outlier point in the optimal solution. 
%To overcome the lack of knowledge about such a point, 
%function \textproc{1-Center} in \Cref{alg:1-center}
%tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and 
%generates a ball for each, using function \textproc{1-Center-Special}. 
%It returns the ball with the minimum radius. 
%Note that, in the streaming model, %due to space limitations inherent to streaming data model, 
%the body of the for loop in function \textproc{1-Center} runs in parallel.
%To run function \textproc{1-Center} in the data stream model,
%we run the body of the for loop in $z+1$ parallel streaming instances.

\begin{theorem} \label{thm:1-center}
	Algorithm~\ref{alg:1-center} computes a $2$-approximation to
	the 1-center problem with $z$ outliers,
	provided that a non-outlier point in the optimal solution in given.
\end{theorem}

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution,
and $c$ be a non-outlier in the optimal solution.
Since $c$ is covered by $B^*$,
for all points $p \in B^*$, we have
$\len{cp} \le \len{cc^*} + \len{c^*p} \le 2r^*$.
Among the $z+1$ points furthest from $c$,
there is at least one point $q$ which is not outlier in the optimal solution,
and therefore, is contained in $B^*$
(see \Cref{fig:1center}). 
Thus, $\len{cq} \le 2r^*$,
and therefore, the ball $B(c, \len{cq})$ returned by Algorithm~\ref{alg:1-center} 
is a 2-approximation. 
\end{proof}

\begin{figure}[t]
	\centering
	\includegraphics[width=12em]{figs/one-center}
	\caption{Proof of \Cref{thm:1-center}}
	\label{fig:1center}
\end{figure}

\noindent
Algorithm~\ref{alg:1-center} 
assumes that a point $c$ is known in advance
which is guaranteed to be a non-outlier in the optimal solution. 
To overcome the lack of knowledge about such a point, 
we run $z+1$ instances of \Cref{alg:1-center} in parallel,
each of which is given as $c$ one of the first $z + 1$ points in the stream.
Clearly, there exists a point among the first $z+1$ points of $P$
which is not an outlier in the optimal solution.
Therefore, the smallest ball among the $z+1$ balls computed in parallel
is always within factor~2 of the optimal solution. % by \Cref{thm:1-center-special}.
Since the space complexity of Algorithm~\ref{alg:1-center} for one instance is $\O(zd)$, 
and its update time is $\O (zd \log z)$, % and the query time is $\O (z)$.
we get the following result.

\begin{theorem} \label{thm:1-center-stream}
	Given a stream of points in $d$ dimensions,
	we can maintain a 2-approximation to the 1-center with $z$ outliers
	in $\O(z^2d)$ space and $\O (z^2d\log z)$ update time.
\end{theorem}

%function \textproc{1-Center} in \Cref{alg:1-center}
%tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and 
%generates a ball for each, using function \textproc{1-Center-Special}. 
%It returns the ball with the minimum radius. 
%Note that, in the streaming model, %due to space limitations inherent to streaming data model, 
%the body of the for loop in function \textproc{1-Center} runs in parallel.
%To run function \textproc{1-Center} in the data stream model,
%we run the body of the for loop in $z+1$ parallel streaming instances.
%
%The space complexity of Algorithm~\ref{alg:1center} is $\O(z^2 + zd)$, the update time is $\O (z \log z + zd)$ and the query time is $\O (z)$.
%If a non-outlier point $p$ is known then the space, update time and query time complexities of Algorithm~\ref{alg:1center} can be reduced by a factor of $z$.

%-------------------------- 2-center ---------------------------------

\section{The 2-Center Problem with Outliers}
\label{sec:2-center}

In this section, we provide a $(1.8 + \eps)$-approximation algorithm
for the 2-center problem with outliers.
In all algorithms presented in this section,
we assume that the first point of the stream, $p_1$, is non-outlier. 
This assumption can be easily removed by considering $z + 1$ parallel instances of the algorithm, 
similar to what we did in \Cref{sec:1-center}.

Let $B_1^*$ and $B_2^*$ be the optimal balls
in an optimal solution to the 2-center problem with $z$ outliers on a point set $P$.
We denote by $r^*$ the optimal radius,
and by $\delta^*$ the distance between $B_1^*$ and $B_2^*$.
To prove our main result, we distinguish between two cases.
In \Cref{subsec:bigger}, we address the case where $\delta^* > \alpha r^*$, 
for some constant $\alpha$ to be fixed later.
(It will be turn out that $\alpha = 14$ is a proper choice.)
We then present in \Cref{subsec:smaller} 
our algorithm for the case of $\delta^* \leq \alpha r^*$.

%We inspect the case where $\delta^* \ge \alpha r^*$ in Subsection~\ref{subsec:bigger}.
%give a description of our algorithm for a given $r > 0$ for the case where $\delta^* < \alpha r^*$ and $1.2r^* \le r < (1.2 + 2\eps/3)r^*$, which returns a $(1.8 + \eps)$-approximate solution. We explain how to find such an $r$ and present a full description of our algorithm for the case where $\delta^* < \alpha r^*$ in Subsection~\ref{subsec:findr}.

% -------------------------------- First Case -----------------------------------

\subsection{The Case $\delta^* > \alpha r^*$}
\label{subsec:bigger}

In this section, we present a $1.8$-approximation algorithm for the case 
where optimal balls are separated by a distance greater than $\alpha r^*$.
%The value of $\alpha$ will be fixed later in this section.

\begin{obs}
\label{obs:c+4}
	Let $B_1$ and $B_2$ be two congruent balls of radius $r$,
	with distance $\delta > \alpha r$.
	%If $p$ and $q$ are two points in $B_1$ and $B_2$, respectively,
	For any two points $p \in B_1$ and $q \in B_2$,
	we have $1 \le \frac{\len{pq}}{\delta} <  \frac{\alpha + 4}{\alpha}$.
\end{obs}

\begin{proof}
The distance between $p$ and $q$ is at most $\delta + 4r$.
%(see \Cref{fig:c+4}).
Hence, $\frac{\len{pq}}{\delta} \leq 1 + {4r \over \delta} < 1 + \frac{4}{\alpha}$.
\end{proof}

\REM{
\begin{figure}[th]
	\centering
	\includegraphics[width=20em]{figs/c_plus_4}
	\vspace{-1em}
	\caption{Illustrating Observation~\ref{obs:c+4}}
	\label{fig:c+4}
\end{figure}
}

\begin{obs}
\label{obs:intersection}
	Let $B_1$ and $B_2$ be two disjoint balls of distance greater than $\delta$,
	and let $B$ be an arbitrary ball of radius at most $\frac{\delta}{2}$. 
	Then $B$ intersects at most one of $B_1$ and $B_2$.
\end{obs}

\noindent 
%Before presenting our algorithm,
We next prove some properties regarding 
the optimal balls, $B_1^*$ and $B_2^*$.

\begin{lemma}
\label{lem:c-sep}
	Let $B_1^*$ and $B_2^*$ be $\alpha$-separated, with $\alpha > 4$.
	If $p$ is a point in $B_1^*$,
	and $S$ is a $(z+1)$-subset of $P$ furthest from~$p$,
	then $S \cap B_2^*$ is non-empty.
\end{lemma}

\begin{proof}
Suppose by way of contradiction that $S \cap B_2^*$ is empty. 
Since $\card{S} = z + 1$, there is at least one point in $S$
which is not outlier, and hence, it is in $B_1^*$.
Let $q$ be the point in $S \cap B_1^*$ furthest from~$p$. 
Consider $B(p, \len{pq})$. It is clear that $B$ covers $P \setminus S$. 
Since $p, q \in B_1^*$, then $\len{pq}$ is at most $2r^*$. 
Thus, by Observation~\ref{obs:intersection}, $B_2^* \cap B = \emptyset$. 
Therefore, $B_2^* \cap P = \emptyset$, and hence,
 $B_2^*$ is empty, which contradicts the optimality of the solution.
\end{proof}

\begin{lemma}
\label{lem:(z+1)-furthest}
	Let $p$ be a point in $B_1^{*}$,
	and $q$ be the $(z+1)$-furthest point from $p$. 
	Then, $\delta^* > \frac{\alpha}{\alpha+4}\len{pq}$.
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:c-sep}, there exists a point $q' \in B_2^*$ 
such that $\len{pq'} \ge \len{pq}$. 
Thus, by Observation~\ref{obs:c+4}, 
%$\frac{\alpha}{\alpha+4} \len{pp_2} \le \delta^*$. 
${\len{pq} \over \delta^*} \le {\len{pq'} \over \delta^*} < \frac{\alpha+4}{\alpha}$.
The claim follows immediately.
\end{proof}

\begin{lemma}
\label{lem:}
	If $p \in B_1^*$ and $q \in B_2^*$,
	then at most $z$ points of $P$ lie outside $B(p, 2r^*) \cup B(q, 2r^*)$.
\end{lemma}

\begin{proof}
	TBA.
\end{proof}

\begin{lemma}
\label{lem:}
	Let $S$ be a subset of $P$ of size at least $\dz$,
	enclosed by a ball of radius $\delta^* / 2$.
	Then the centerpoint of $S$ lies either inside $B_1^{*}$ or inside $B_2^{*}$.
\end{lemma}

\begin{proof}
	TBA.
\end{proof}

\paragraph{The Algorithm}
We now describe our algorithm for handling the case  $\delta^* > \alpha r^*$.
At any time, our algorithm maintains a partition of $P$ into three disjoint subsets
$B_1$, $B_2$, and Buffer.
The first point $p_1$ is assumed, w.l.o.g, to be in $B_1^*$. 
The algorithm tries to partition points in such a way that 
at the end,
$B_1$ contains the whole $B_1^*$, and $B_2$ contains the whole $B_2^*$,
with possibly some outliers being contained in $B_1$ and $B_2$.
%Buffer has always size at most $z$.
The algorithm sets $c_1 = p_1$ as the fixed center of $B_1$,
and picks $c_2$ among the points processed so far as a candidate center for $B_2$.
Moreover, the algorithm maintains two values $\delta$ and $r$,
where at any time, $\delta$ is a lower bound of $\delta^*$, 
and $r$ is an upper bound of $2r^*$.

Our algorithm is presented in \Cref{alg:case-1}.
For each input point $p \in P$, the algorithm first tries 
to add $p$ to either $B_1$ or $B_2$,
using functions \textproc{AddToB$_1$} and \textproc{AddToB$_2$}, respectively.
%presented in Algorithms~\ref{alg:addB1} and~\ref{alg:addB2}.
If none of them fits, the point is added to Buffer.
The function \textproc{AddToB$_1$} adds a point $p$ to $B_1$
only if it is within distance $\delta$ of the center $c_1$.
Similarly, \textproc{AddToB$_2$} adds a point $p$ to $B_2$
only if it is within $\rc$-radius of $c_2$. 
The two functions also update $\delta$ and $\rc$ whenever necessary,
to maintain the invariants to be defined shortly in Lemma~\ref{lem:invariants}.

\begin{algorithm}[t]
\caption{\sc 2-CenterCase1$(P)$} 
\algsetup{indent=1.5em}
\label{alg:case-1}
\begin{algorithmic}[1]
\STATE $c_1 \gets p_1$, $\rc  \gets 0$, $\delta \gets 0$
\FORALL{$p \in P$}
	\IF {\NOT (\Call{AddToB$_1$}{$p$} \OR \Call{AddToB$_2$}{$p$})}
		\STATE add $p$ to Buffer
		\WHILE{$\card{\mbox{Buffer}} > z$} \label{step:buffer-overflow}
			\IF{$\card{B_2} \ge \dz$}
				\STATE $B_1 \gets B_1 \cup B_2$, $B_2 \gets \emptyset$
			\ELSIF{$c_2$ is set}
				\STATE $B_1 \gets B_1 \cup \set{c_2}$ %, $B_u \gets B_u \cup \set{c_2}$
				%\STATE remove $c_2$ from $B_2$
			\ENDIF
			\STATE $T \gets \mbox{Buffer} \cup B_2 \setminus \set{c_2}$
			\STATE $B_2 \gets \emptyset$ 
%			\IF{first iteration of while}
%				\FOR{$p \in T$}
%					\IF{\Call{AddToB$_1$}{$p$}}
%						\STATE remove $p$ from $T$
%					\ENDIF
%				\ENDFOR
%			\ENDIF
			\STATE $c_2 \gets$ an arbitrary point of $T$
			\STATE $\rc  \gets \frac{2}{\alpha}\len{c_1 c_2}$
			\FOR{$p \in T$}
				\STATE \Call{AddToB$_2$}{$p$}
			\ENDFOR				
			\STATE $\mbox{Buffer} \gets T \setminus B_2$
		\ENDWHILE
	\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht]
\caption{\sc AddToB$_1(p)$} 
\algsetup{indent=1.5em}
\label{alg:addB1}
\begin{algorithmic}[1]

	%\STATE $q \gets c_1$
	\IF {at least $z + 1$ points have been arrived}
		\STATE $q \gets$  $(z+1)$-furthest point from $c_1$
	\ELSE
		\STATE $q \gets c_1$
	\ENDIF
	\STATE $\delta \gets \frac{\alpha}{\alpha+4} \len{c_1q}$
	\IF {$p \in B(c_1, \delta)$}
		\STATE $B_1 = B_1 \cup \set{p}$ %, B_u = B_u \cup \set{p}$
		\RETURN{true}
	\ENDIF
	\RETURN{false}

\end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht]
\caption{\sc AddToB$_2(p)$} 
\algsetup{indent=1.5em}
\label{alg:addB2}
\begin{algorithmic}[1]
%\FUNCTION{AddToB$_2$}{p}
	\IF{$c_2$ is set \AND $p \in B(c_2, \rc )$}
		\STATE $B_2 \gets B_2 \cup \set{p}$ %, B_u \gets B_u \cup \set{p}$
		\IF{$|B_2| = \dz$}
			%\STATE $\cp  \gets$  centerpoint of $B_2$
			\STATE $\rc  \gets  (2 + \frac{2}{\alpha}) \times \rc$
			\FOR{$p$ in Buffer}
				\IF{$p \in B(c_2, \rc )$}
					\STATE $B_2 \gets B_2 \cup \set{p}$ %, B_u \gets B_u \cup \set{p}$
					\STATE remove $p$ from Buffer
				\ENDIF
			\ENDFOR
		\ENDIF
		\RETURN{true}
	\ENDIF
	\RETURN{false}
%\ENDFUNCTION
\end{algorithmic}
\end{algorithm}

Whenever the buffer overflows (in line~\ref{step:buffer-overflow} of \Cref{alg:case-1}),
%the assumption about $\cp \in B_2^*$ is contradicted.
%At some point during the execution of Algorithm~\ref{alg:case-1}, the assumption about $\cp \in B_2^*$ might be contradicted when the buffer overflows. We can then conclude that
% $\cp \notin B_2^*$.
the algorithm takes one of the following actions
depending on the size of $B_2$. 
If $|B_2| \ge \dz$, then the points of $B_2$ are moved to $B_1$, and $B_2$ is reset.
Otherwise, the old $c_2$ (if already set) is moved to $B_1$,
and another point from $T = B_2 \cup \Buffer \setminus \set{c_2}$ is picked as $c_2$.
The while loop iterates $O(dz)$ times,
because after the first iteration, we are sure 
that $T$ has at most $\dz + z$ points, from which 
one point (i.e., $c_2$) is removed at each subsequent iteration.

\paragraph{Analysis}
For the sake of analysis, we maintain a special point,
denoted by $\cp$, %not necessarily in $P$,
which we call a \emph{central point}.
The point is defined as follows:
if $\card{B_2} < \dz$, then $\cp = c_2$, otherwise,
$\cp$ is the centerpoint of the first $\dz$ points currently in $B_2$.

\begin{lemma}
\label{lem:invariants}
	The following invariants are maintained during the execution of the algorithm:

\begin{enumerate}
\item [(a)] $\delta < \delta^*$ %$\delta$ is a lower bound of $\delta^*$
\item [(b)] $B_1 \cap B_2^* = \emptyset$
\item [(c)]  if $\cp \in B_2^*$, then 
	\begin{enumerate}
		\item [1.] $2r^* \le r \le \delta /2$
		\item [2.] $B_2 \cap B_1^* = \emptyset$
		\item [3.] all points in $\Buffer$ are outliers
	\end{enumerate}
\end{enumerate}
\end{lemma}

\begin{proof}
TBA \ 
(All non-used parts from the previous version are copied here.
They might be useful for composing the proof of this lemma.)

The function \textproc{AddToB$_1$} presented in \Cref{alg:addB1} 
is used to add a point $p$ to $B_1$. 
%The point $c_1$ is the first point of the stream,
%which is assumed to be non-outlier, and is in $B_1^*$.
The point will be added to $B_1$ only if it is within distance $\delta$ of the center $c_1$, 
where $\delta = \frac{\alpha}{\alpha+4}\len{qc_1}$ 
and $q$ is the $(z+1)$-furthest point from $c_1$.
%among the points of $P$ seen so far.
By Lemma~\ref{lem:(z+1)-furthest}, $\delta$ is a strict lower bound of $\delta^*$. 
Therefore, any point within $\delta$-radius of $c_1$ is guaranteed to be outside $B_2^*$. 
Thus, property P1 is not violated by this function.
%$B_1 \cap B_2^* = \emptyset$.
Note that if less than $z+1$ points have been seen so far,
then $\delta = 0$, %$q = c_1$, 
and hence, property P1 is still maintained.

The function \textproc{AddToB$_2$} presented in \Cref{alg:addB2}
is used to add a point $p$ to $B_2$.
Initially, $B_2$ is empty and $c_2$ is not initialized. 
When an appropriate candidate is found, 
$c_2$ is set to that candidate by the main algorithm (Algorithm~\ref{alg:case-1}),
and $\rc$ is set to $\frac{2}{\alpha}\len{c_1c_2}$.
After that, points are added to $B_2$ by function \textproc{AddToB$_2$}
only if they are within $\rc$-radius of $c_2$. 
Consider $c_2$ as the central point $c_p$. % is considered as the central point $\cp$. 
Clearly, property P2 holds initially when $B_2$ consists of a single point $c_2$. 
All subsequent points added to $B_2$ are within $\rc$-radius of $c_2$.
Therefore, for $\alpha > 7$ (Why?), Observation~\ref{obs:intersection}
implies that $B_2 \cap B_1^* = \emptyset$,
and hence, property P2 is maintained. %invariant under this operation.

When the size of $B_2$ reaches $\dz$, 
we update the central point $\cp$ to the centerpoint of $B_2$.
%We claim that $B(c_2, (2 + \frac{2}{\alpha})\rc )$ completely covers
%the ball $B(\cp, \frac{2\len{c_1\cp}}{\alpha})$.
If $\cp \in B_2^*$, then by Observation~\ref{obs:c+4},
we have $\frac{\len{c_1\cp}}{\delta^*} \geq 1$,
and hence, $r^* < \frac{\len{c_1\cp}}{\alpha}$. 
Therefore, $B_2^* \subset B(\cp, \frac{2\len{c_1\cp}}{\alpha})$.
Moreover,
\[
	2 \cdot \frac{\len{c_1\cp}}{\alpha} \leq \frac{2}{\alpha}(\len{c_1c_2} + \rc ) 
	\leq (1 + \frac{2}{\alpha})\rc.
\]
Therefore, $B_2^* \subseteq B(c_p, (1 + \frac{2}{\alpha})\rc )
\subseteq B(c_2, (2 + \frac{2}{\alpha})\rc )$. 
It is easy to see that $(2 + \frac{2}{\alpha})\rc  \le \frac{\delta}{2}$ provided that $\alpha \ge 14$.
Thus, if we increase $\rc$ by a factor of $(2 + \frac{2}{\alpha})$, 
%by setting $B_2 = B(c_2, (2 + \frac{2}{\alpha})\rc )$, 
$B_2 \cap B_1^* = \emptyset$
by Observation~\ref{obs:intersection}, and hence,
property P2 is still maintained.

Whenever the buffer overflows 
%the assumption about $\cp \in B_2^*$ is contradicted.
%At some point during the execution of Algorithm~\ref{alg:case-1}, the assumption about $\cp \in B_2^*$ might be contradicted when the buffer overflows. We can then conclude that
% $\cp \notin B_2^*$.
the algorithm takes one of the following actions
depending on the size of $B_2$.

\begin{itemize}
\item 
If $|B_2| < \dz$, then $c_2$ can be added to $B_1$
without violating P1. (Why?) 
The algorithm then tries the points in $B_2 \cup \mbox{Buffer}$, 
until it finds a suitable candidate for $c_2$ that does not leave
more than $z$ points uncovered. 
Note that at most $(z+1)(d+1) + z$ points have to be tried.

\item 
If $\card{B_2} \ge \dz$, let $q$ be the $(z+1)$-furthest point from $p_1$,
and let $p'$ be the nearest point of the ball $B_2(c_2, \rc )$ to $p_1$. 
Then $\len{qp_1} > \len{p'p_1}$, since $B_2$ contains more than $z+1$ points. 
Therefore, by Lemma~\ref{lem:(z+1)-furthest} we have
\[
\delta \ge \frac{\alpha }{\alpha+4}\len{qp_1} \ge \frac{\alpha}{\alpha+4} (\len{c_1c_2} - \rc ) = \frac{\alpha^2 - 2\alpha}{2\alpha + 8} \rc.
\]
As a result, if $\alpha \geq 14$ then $(2 + \frac{2}{\alpha}) \rc  < \frac{\delta}{2}$. 
By Observation~\ref{obs:intersection}, the ball $B_2(c_2, (2 + \frac{2}{\alpha}) \rc )$ 
has nonempty intersection with exactly one of $B_1^*$ and $B_2^*$ 
(i.e., not all of its points can be outliers). 
Since the buffer overflowed, the assumption about the central point, 
$\cp$, being in $B_2^*$ has become invalid. 
By Observation~\ref{obs:omitting-centerpoint}, if $\cp$ is not covered, 
then the ball $B_1^*$ leaves at least $z+1$ points of $B_2$ uncovered, 
and there will be more than $z$ outliers. 
Thus, $\cp$ has to be covered. 
As a result, $\cp \in B_1^*$ and all the points of $B_2$ can be added to $B_1$. 
This is done by replacing $B_1$ with $B_1 \cup B_2$, and resetting $B_2$ to empty set.
\end{itemize}
\end{proof}

\begin{lemma}
\label{lem:}
	After termination of \Cref{alg:case-1}, 
	if $|B_2| \ge \dz$ and $\cp \not\in B_2^*$,
	then $B_2 \cap B_2^* = \emptyset$.
\end{lemma}

\begin{proof}
	TBA.
\end{proof}

\paragraph{Streaming Model}
To implement our algorithm in the streaming model,
we maintain the sets $B_1$ and $B_2$ 
in a data structure that supports adding points,
and gives a $\beta$-approximation to 1-center with $k$ outliers, for $k=0,\dots,z$.
Moreover, we maintain a set $B_u = B_1 \cup B_2$ in a similar data structure.
Note that these data structures do not need to maintain all the points. 
They only need to have a buffer of size $\dz$ that keeps the most recently added points.

Let $B_1'$, $B_2'$, and $B_u'$ be the $\beta$-approximate balls obtained from the data structures
[This part needs clarification].
%1-centers obtained from $B_1$, $B_2$ and $B_u$ by the $\beta$-approximation algorithm, respectively. 
To answer a query, we can use these balls as follows. 
By our initial assumption about $p_1$, $B_1$ contains $B_1^*$,
and $B_1'$ is a $\beta$-approximation for $B_1^*$. (Why?)
But for $B_2$, it might be the case that our assumption about $c_2$ 
was incorrect, so $B_2'$ may not be a good approximation for $B_2^*$. 
There are two cases to be considered based oon $\card{B_2}$. 
If $\card{B_2} < (z+1)(d+1)$, then we can try each point in 
$B_2 \cup \mbox{Buffer}$ as a candidate for the central point. 
If $\card{B_2} \ge \dz$, then there is no need for points in $B_2$ to be considered, 
since for those points as central point, $B_1$ will be replaced by $B_u$ and $B_2$ will be empty. (Why?)
Thus it suffices to consider points of Buffer as candidates 
and compare them to the solution of $(B_u, \emptyset)$.

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
%\Function{Query}{}
	\STATE solutions $\gets [\Call{ComputeMin}{B_u, \emptyset, \text{Buffer}}]$

	\STATE candidates $\gets$ Buffer
	\IF {$\card{B_2} < \dz$}
		\STATE candidates $\gets \text{Buffer} \cup B_2$ % \setminus \set{c_2}$
	\ENDIF

	\FOR{$c_2 \in$ candidates}
		\STATE $\rc  \gets \frac{2}{\alpha} \len{c_1 c_2}$
		\STATE $B_2 \gets \emptyset$, Buffer $\gets \emptyset$ 
		\FOR{$p \in$ candidates}
			\IF{\NOT \Call{AddToB$_2$}{$p$}}
				\STATE add $p$ to Buffer
			\ENDIF
		\ENDFOR
		\STATE add \Call{ComputeMin}{$B_1$, $B_2$, Buffer} to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\sc ComputeMin($B_1, B_2, \text{\rm Buffer}$)} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
%\Function{ComputeMin}{$B_1$, $B_2$, Buffer}
	\STATE solutions $\gets [\,]$
	\FOR{$k \gets 0, \dots,(z - \card{\mbox{Buffer}})$}
		\STATE $r_1 \gets \Call{1-Center}{B_1, k}$
		\STATE $r_2 \gets \Call{1-Center}{B_2, z - \card{\mbox{Buffer}} - k}$
		\STATE add $\max \set{r_1, r_2}$ to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

\REM{
Suppose that our data structure for maintaining $B_1$, $B_2$ and $B_u$ uses $S(n,z,d)$ space, $T(n,z,d)$ update and $Q(n,z,d)$ query time. Note that by $Q_{o}(zd,z,d)$ we mean the query time for the offline 1-center problem with $z$ outlier in $d$ dimensions. The best known algorithm for 1-center without outliers is $1.22$-approximation by~\cite{chan2014streaming, agarwal2010streaming}. The buffering framework introduced in~\cite{zarrabi2009streaming} with $\O(dz^3 + (\frac{d}{\eps^3})\log(\frac{1}{\eps}))$ space complexity, can be used with this algorithm as subroutine to achieve a $(1.22\sqrt{2})$-approximation ($1.22\sqrt{2} < 1.8$) for 1-center problem with outliers.

\begin{theorem}
\label{thm:2nd-cmplx}
Space complexity of Algorithm~\ref{alg:case-1} is $\O(dz+S(n,z,d))$. It takes $\O^*(dzT(n,z,d))$ time for update. Query time in Algorithm~\ref{alg:query} is $\O(zQ(n,z,d) + dz(dz+zQ_o(zd,z,d)))$ and it returns two centers that guarantee a $(1.8 + \eps)$-approximation for 2-center problem with outliers.
\end{theorem}

\begin{proof}
The space complexity for this algorithm is derived from the space used by $B_1$, $B_2$, $B_u$ and the Buffer which is obviously $\O(dz+S(n,z,d))$. The while loop runs at most once for each point, so amortized time is $\O(zdT(n,z,d))$. The candidates in the Algorithm~\ref{alg:query} are at most $zd$ points  so the query time will easily follows.
\end{proof}
}

\begin{theorem}
\label{thm:far-cmplx}
	If $\delta^* > \alpha r^*$, a $1.8$-approximation to the 2-center problem with $z$ outliers 
	can be maintained in $O(d^3z^2)$ space and $\poly(d,z)$ update time.
\end{theorem}

\begin{proof}
The streaming algorithm of Zarrabi-Zadeh and Mukhopadhyay~\cite{zarrabi2009streaming} 
equipped with the streaming $1.22$-approximation for 
1-center~\cite{agarwal2010streaming, chan2014streaming}
yields a $(1.22 \times \sqrt{2})$-approximation algorithm
for 1-center with $z$ outliers.
Since $1.22 \times \sqrt{2} < 1.8$, the approximation factor of our algorithm follows.
The algorithm of~\cite{zarrabi2009streaming}
has $O(d^3z)$ space and $\poly(d,z)$ update time,
which implies the same bounds for \Cref{alg:case-1}.
Since we need to run $z + 1$ instances of \Cref{alg:case-1} in parallel, 
the space and update time is increased by a factor of $O(z)$.
\end{proof}

% -------------------------------- Second Case -----------------------------------

\subsection{The Case $\delta^* \leq \alpha r^*$}
\label{subsec:smaller}
Our idea in this section is derived from~\cite{ahn2014computing, kim2014improved}. To avoid duplication, we just outline the important parts of their algorithm and our modifications to it. Kim's algorithm has $10$ different states. In each step one or many of the states can be valid and their algorithm considers all of them in parallel. In each state, there exist at most two balls. A transition between the states occurs whenever a point not covered by any of the two balls has arrived in the stream. This transition graph starts with a blank node (with no balls) and has depth four (without considering the end nodes). It is straightforward to see that this transition graph is a DAG.
Our modification is on the transition part.
Points that are covered by current solution can safely be ignored, as they do not cause any transitions in the state, and only those points that lie outside the solution are candidates for being an outlier.
In each state the number of points that should be considered as outliers is unknown. So all the possible choices should be considered. It is sufficient to have four integers, $n_1,\dots, n_4$ representing the number of outliers in depth $1$ to $4$, such that $\sum_{i=1}^{4} n_i=z$. 
The modified algorithm is presented in \Cref{alg:smallc}.
In this algorithm, \textproc{NthOutlier}$(n,p)$ returns the $n$-th point 
arriving after $p$ in $P$ that does not lie in the current candidate solution.

\begin{algorithm}
\caption{\sc 2-CenterCase2$(P, z, r)$} 
\algsetup{indent=1em}
\label{alg:smallc}
%\leavevmode
\begin{algorithmic}[1]
%\Function{Known$r$}{$P, z, r$}
 	\STATE list $\gets [\,]$
	\FORALL{$(n_1, \dots, n_4)$ such that $\sum n_i=z$}
		\STATE $CS \gets B(p_1,r)$
		\STATE $p^o_{2} \gets $ \Call{NthOutlier}{$n_1 + 1$, $p_1$}
		\IF{$p^o_{2}$ arrives}
			\STATE $CS\gets [ \text{subcase $p^o_{2} \in B_1^*$} ] \cup [ \text{subcase $p^o_{2} \in B_2^*$} ]$
			\FORALL{case $S$}
				\STATE $p^o_{3} \gets$ \Call{NthOutlier}{$n_2 + 1$, $p^o_{2}$}
				\IF{$p^o_{3}$ arrives} 
					\STATE $CS\gets CS\setminus S \cup [ \text{subcase $p^o_{3} \in B_1^*$} ] \cup [ \text{the subcase $p^o_{3} \in B_2^*$}]$
					\FORALL {subcase $S'$} 
						\STATE $p^o_{4} \gets$ \Call{NthOutlier}{$n_3 + 1$, $p^o_{3}$}
						\IF {$p^o_{4}$ arrives} 
							\STATE replace $S'$ by a new candidate solution
							\STATE $p^o_{5} \gets$ \Call{NthOutlier}{$n_4 + 1$, $p^o_{4}$}
							\IF{$p^o_{5}$ arrives}
								\STATE abandon the solution
							\ENDIF
						\ENDIF
					\ENDFOR
				\ENDIF
			\ENDFOR

		\ENDIF
		%\STATE $\mathop{Sol}_{n_1,n_2,n_3,n_4} \gets$ the solution with the smallest larger radius, among all the subcases.
		\STATE add to list the solution with the smallest larger radius, among all subcases
	\ENDFOR
	\RETURN the solution in list with the smallest radius
\end{algorithmic}
\end{algorithm}

\begin{theorem}
	For $1.2r^* \le r < (1.2 + 2\eps/3)r^*$ and $\delta^* < \alpha r^*$,
	Algorithm~\ref{alg:smallc} returns a $(1.8+\eps)$-approximate solution,
	using $\O (dz^3)$ space and $\O (dz^3)$ update and query time.
\end{theorem}

\begin{proof}
Since our algorithm considers all possible input cases of streaming points, there is at least one feasible solution. Since every feasible solution has its larger radius at most $3r/2$, the final solution has larger radius at most $3r/2 \le (1.8 + \eps)r^*$.
For space complexity, our algorithm maintains at most two balls in each case, and therefore it uses $\O(dz^3)$ space. Whenever the next point is inserted, the algorithm updates the solution for each subcase in $\O (d)$ time. Therefore, the algorithm spends $\O (dz^3)$ update time for each point of $P$. Answering a query consists of choosing the minimum radius among all the candidate solutions, which amounts to $\O (dz^3)$ time.
\end{proof}

\subsubsection{Finding $r$}
\label{subsec:findr}

\begin{lemma}
\label{lem:1-2}
	An optimal solution for the 1-center problem with $z$ outliers on a point set
	gives a $(2 + {\alpha \over 2})$-approximation for
	the 2-center problem with $z$ outliers on that set,
	provided that $\delta^* \le \alpha r^*$.
\end{lemma}

%\begin{lemma}
%\label{lem:1-2}
%	Let $r_1^*$ and $r^*$ be optimal solutions for
%	the 1-center and 2-center problems with $z$ outliers, respectively.
%	If the optimal 2-center balls have distance at most $\alpha r^*$, 
%	then $r^* \le r_1^* \le \left(2 + \frac{\alpha}{2}\right) r^*$.
%\end{lemma}

\begin{proof}
Let $r_1^*$ and $r^*$ be the optimal radii for
the 1-center and 2-center problems with $z$ outliers on $P$, respectively.
It is clear that $r^* \le r_1^*$,
because any feasible solution $B^*$ for 1-center with $z$ outliers
yields a feasible solution $(B^*, B^*)$ for 2-center with $z$ outliers.
%Let $B^*$ be an optimal solution for the 1-center problem with $z$ outliers. 
%Then, $(B^*, B^*)$ is a feasible solution for the 2-center problem with $z$ outliers. 
%Hence, $r_2^* \le r_1^*$.
Now, suppose that $B_1^*(c_1^*, r^*)$ and $B_2^*(c_2^*, r^*)$
are the balls in an optimal solution for the 2-center problem with $z$ outliers.
Let $c$ be the midpoint of the segment connecting 
$c_1^*$ to $c_2^*$ (see Figure~\ref{fig:2lt1}).
Clearly, $B\left(c, \frac{\delta}{2} + 2r_2^*\right)$ covers both $B_1^*$ and $B_2^*$. 
Therefore, it is a feasible solution for the 1-center problem with $z$ outliers.
Hence, $ r_1^* \le \left(2 + \frac{\alpha}{2}\right) 2^*$.
\end{proof}

\begin{figure}[th]
	\centering
	\includegraphics[width=19em]{figs/2lt1}
	\caption{Proof of Lemma~\ref{lem:1-2}}
	\label{fig:2lt1}
\end{figure}

\noindent 
The following is a direct corollary of Lemma~\ref{lem:1-2} and Theorem~\ref{thm:1-center}.

\begin{cor}
	If $\delta^* \le \alpha r^*$, 
	\Cref{alg:1-center} computes a $(4 + \alpha)$-approximation to $r^*$. 
\end{cor}

\begin{obs}
\label{obs:2apr}
Let $r_1$ and $k_0$ be two positive real numbers. Define $k=k_0 2^i$ to be the smallest real number satisfying the inequality $k \ge r_1$, where $i$ is a non-negative integer. Clearly, $k$ is a $2$-approximation for $r_1$.
\end{obs}

Our algorithm maintains $m=\ceil{{1.2 \times (3 \alpha+12)}/{\eps}}$ candidate lengths.
For each candidate length $t$, assume that $r = t$ and use it to run Algorithm~\ref{alg:smallc}. Algorithm~\ref{alg:1-center} is used to estimate $r$. Let $k_i$ be the $i$-th non-zero radius calculated by the algorithm. Obviously, the sequence of answers given by the algorithm is increasing.

Let $l_1=k_1$ and $l_i=2^j l_1$, where $l_i$ is the smallest number satisfying $l_i \ge k_i$. By Observation~\ref{obs:2apr}, $l_i$ is a $(8 + 2\alpha)$-approximation to $r_2^*$.
If the interval $(0, 1.2 l_i]$ divides into $m$ equal segments (Figure~\ref{fig:findr}), then length of each segment is at most $2/3 \eps r_2^*$ and the endpoints as candidates are $L_i = \set{j \times t_i \provided j = 1, \dots, m}$ where $t_i = \frac{1.2 l_i}{m}$. 
The inequality $1.2 r_2^* \le j \times l_i \le (1.2 + \frac{2\eps}{3})r_2^*$ holds for at least one of these candidates.

\begin{figure}[h]
	\centering
	\includegraphics[width=21em]{figs/findr}
	\caption{Subdivision of the interval $(0, 1.2 \ell_i]$}
	\label{fig:findr}
\end{figure}

When the first non-zero radius is observed, $l_1$ is set. Algorithm~\ref{alg:smallc} should be executed for each candidate in $L_1$ for all the previous points, which have been stored in a buffer. It is not hard to see that the space complexity for storing these points is $\O(z)$. For each new point, if $l_i=l_{i-1}$ then $L_i = L_{i-1}$ and it suffices to add the new point to all parallel instances corresponding to candidates in $L_i$. Otherwise, Algorithm~\ref{alg:smallc} should be executed for all of the candidates in $L_i$, with all the points. But this is not achievable due to space limitations. To overcome this problem, note that if the candidate $x_j=j \times t_i \in L_i$ is also in $L_{i-1}$, there is no need to start Algorithm~\ref{alg:smallc} over and the current execution will be continued. If $x_j \notin L_{i-1}$, then it is certain that $x_j \ge l_{i - 1}$ (Observe that $l_i=ul_{i-1}$, for some $u\in \mathbb{N}$). Since the points that do not lie in the candidate solution are saved in the buffer used in Algorithm~\ref{alg:1-center} (which has size at most $z$), all the non-outlier points of this algorithm lie in the candidate balls in Algorithm~\ref{alg:smallc}. These outliers have been stored in a buffer, so for the new candidate length, it suffices to execute Algorithm~\ref{alg:smallc} with only the outlier points of $B_1$.

%\begin{theorem}
%\label{thm:1st-cmplx}
%Algorithm~\ref{alg:smallc} uses $\O(\frac{dz^3}{\eps})$ space, and takes $\O(\frac{dz^4}{\eps})$ and $\O(\frac{dz^3}{\eps})$ time for update and query, respectively.
%\end{theorem}

\begin{theorem}
\label{thm:close-cmplx}
	If $\delta^* \leq \alpha r^*$, a $(1.8 + \eps)$-approximation to the 2-center problem with $z$ outliers 
	can be maintained in $\O(\frac{dz^4}{\eps})$ space 
	and $\O(\frac{dz^3}{\eps})$ update/query time.
\end{theorem}

\subsection{Final Result}

\begin{theorem}
	A $(1.8 + \eps)$-approximation to the 2-center problem with $z$ outliers
	can be maintained in the streaming model using 
	$\O(dz^2 (d^2 + z^2/\eps))$ space and 
	$\poly(d, z, {1 \over \eps})$ update/query time.
\end{theorem}

\begin{proof}
This is immediate from Theorems~\ref{thm:far-cmplx} and~\ref{thm:close-cmplx}.
%and the fact that $p_1$ was assumed to be a non-outlier.
\end{proof}

%\begin{theorem}
%Our algorithm uses $\O(zS(n, z, d) + \frac{dz^4}{\eps})$ space and has time complexity $\O^*(dz^2T(n,z,d) + \frac{dz^5}{\eps})$ and $\O(z^2Q(n,z,d) + dz^2(dz+zQ_o(zd,z,d)) + \frac{dz^4}{\eps})$ for update and query operations, respectively. It returns a $(1.8 + \eps)$-approximate solution for 2-center problem with outliers in any dimension.
%\end{theorem}
%
%\noindent 
%Using the algorithm of~\cite{zarrabi2009streaming}, 
%the space complexity of our algorithm is $\O(\frac{dz^4}{\eps} + \frac{dz}{\eps^3})$.

%---------------------------- Conclusions ---------------------

\section{Conclusions}
\label{sec:conc}

In this paper, we presented a $(1.8 + \eps)$-approximation streaming algorithm for 2-center problem with outliers in Euclidean space.
It improves the previous $(4+\eps)$-approximation algorithm available for the problem 
due to Charikar~\etal~\cite{mccutchen2008streaming}. 
Finding better approximation factor or space complexity
is an interesting problem that remains open.
It is also interesting to see if the ideas in this paper 
can be extended to the $k$-center problem with outliers in the data stream model
for any $k$, or even for small values of $k \ge 3$.

\paragraph{Acknowledgement}
The authors would like to thank Kiana Ehsani and Sahand Mozaffari 
for their thoughtful discussions, 
and for their contribution to improve the presentation of this paper.

%--------------- BIBS --------------------

\small
\baselineskip=.85\baselineskip
\bibliographystyle{abbrv}
\bibliography{bibs/abbrv,bibs/ref}

\end{document}